\prefacesection{Acknowledgments}

First and foremost, I would like to express my sincere gratitude to my advisor, Professor Hamdi Tchelepi, whose neverending encouragement, inspiration and support were vital to the completion of this work.   I highly appreciate his guidance and advice both on academic matters and beyond, as well as his patience and kindness as I faced some obstacles in my research journey.   I have been very fortunate to have him as my research supervisor.

I am very grateful to Dr. Nicola Castelletto for our close collaboration and his mentorship throughout the years.   Our frequent discussions and hours of brainstorming ideas together have contributed immensely to this work, as well as to my own growth and development as a researcher.   I have learned a great deal from him, from overall approach to problems, to the many tips and tricks for effective research, writing and presentation of ideas.   This work could not have been completed without his guidance, and his careful reading of a draft of this dissertation helped improve it in many ways.

I would like to thank Professor Daniel Tartakovsky for being a part of the dissertation reading committee, and Professor Louis Durlofsky for serving on my oral defense as an examiner.   In addition, I acknowledge and thank Professor Eric Darve for kindly chairing the defense.   Their insightful comments and questions lead to an interesting discussion about future extensions of this work and are highly appreciated.

A special thanks goes to the many collaborators from Stanford, Lawrence Livermore National Laboratory and TotalEnergies with whom I've had the pleasure of interacting while working on GEOSX for the past 5 years: Randolph Settgast, Joshua White, Benjamin Corbett, Fran{\c c}ois Hamon, Matteo Cusini, Thomas Gazzola, Herve Gross, Antoine Mazuyer, Jacques Franc, Andrea Franceschini and many others.   Our regular meetings, design discussions and code reviews were invaluable to my growth as a researcher and developer.   

I would like to thank the many friends and colleagues in SUETRI-B (former SUPRI-B) research group as well as the faculty, staff, students and scholars in the Department of Energy Science and Engineering.   I've learned a lot about various aspects of subsurface modeling, the broader energy industry and computational science through our interactions.   Special thanks goes to Dr. Sebastian Bosma and Dr. Olav M{\o}yner for our collaboration on multiscale methods; to Dr. Pavel Tomin, Dr. Timur Garipov and Dr. Igor Shovkun for our many fruitful discussions; to Ruslan Rin, Karine Levonyan, Christine Maier, Youssef Elkady, Ouassim Khebzegga, Jose Eduardo Ramirez for all our time and conversations together.   I would also like to express my gratitude to Professor Denis Voskov for his support early on in my graduate career.

I wish to thank LLNL collaborators and Stanford center for Computational Earth and Environmental Sciences (CEES) for access to computational resources used in this work.   Financial support from industrial affiliates of Stanford's Reservoir Simulation (SUETRI-B) program and particularly from TotalEnergies through STEMS and FC-MAELSTROM projects is gratefully acknowledged.   I would also like to thank TotalEnergies for the many internships opportunities that enriched my experience in many ways.

Finally, my deepest gratitude goes to my parents, Elena and Vladimir, and my wife, Dr. Yulia Makerova, for their unwavering support and encouragement during this most difficult chapter of my life.   Without them, my PhD journey would not have been possible.